Modal Abstractions for Virtualizing Memory Addresses

### 2025-07-19 syntaxsemantics.tex
synsem check

To develop our core logical ideas, we instantiate \iris with a simple language for streams of instructions, 
and a logical machine model corresponding to execution of x86-64 assembly instructions with virtual memory enabled on the 

@@

A register identifier, $\reg$, is chosen from a fixed finite set of register identifiers, $\regset$.\footnote{\add{$\reg$ abbreviates \emph{general register}, which includes integer registers (e.g., \lstinline|rax|),

@@

The most important instructions we model are memory-accessing variants of the x86-64 \lstinline|mov| instruction, which we format in Intel syntax
(destination on the left, source on the right).
Thus a store to memory is $\textsf{mov}~[r_m]~r_r$, and a load from memory is $\textsf{mov}~r_r~[r_m]$.
% \[
% \begin{array}{l}
%   \hbox{(\TirNameStyle{WriteToMemFromReg})} \qquad
%   \textsf{mov}~[r_m]~r_r \\
%   \hbox{(\TirNameStyle{WriteToRegFromMem})} \qquad 
%   \textsf{mov}~r_r~[r_m]
%   \end{array}
%   \]
Operationally, each first translates the virtual address stored in register $r_m$ to a physical address \add{(a page table walk)},
then either updates the memory at that physical location with the contents of $r_r$ (for store)
or updates $r_r$ with the contents of that memory (for load).
% The semantics of each is realized by first translating the virtual-memory address of a memory location stored in $r_m$ to a physical memory location
% per the description in Section \ref{sec:backgroundonmachinemodel},
% then either updating those memory contents with the contents of register $r_r$ (\TirNameStyle{WriteToMemFromReg}) 
% or loading the value at that physical memory location into the register $r_r$
% (\TirNameStyle{WriteToRegFromMem}). 
Our formalization includes
additional \lstinline|mov| variants (e.g., accessing memory at constant offsets from the base register, or moves between registers),
basic integer and bitwise operations (\lstinline|add|,
\lstinline|and|, bit shifts, etc.) with their effects on \lstinline|rflags|, jumps and some (not all) conditional jumps, \lstinline|call|, \lstinline|ret|, \lstinline|push|,
and \lstinline|pop|.
The semantics model updates to the instruction pointer, though we elide that register (\lstinline|rip|)
from specifications in the paper for clarity.
\looseness=-1

>>>>

To develop our core logical ideas, we instantiate \iris with a simple language for a stream of instructions
and a logical machine model corresponding to the execution of x86-64 assembly instructions with virtual memory enabled on the

@@

A register identifier, $\reg$, is chosen from a fixed finite set of register identifiers, $\regset$. \footnote{\add{$\reg$ abbreviates \emph{general register}, which includes integer registers (e.g., \lstinline|rax|),

@@

The most important instructions that we model are memory-accessing variants of the x86-64 \lstinline|mov| instruction, which we format in Intel syntax
(destination on the left, source on the right).
Thus, a store to the memory is $\textsf{mov}~[r_m]~r_r$, and a load from memory is $\textsf{mov}~r_r~[r_m]$.
% \[
% \begin{array}{l}
%   \hbox{(\TirNameStyle{WriteToMemFromReg})} \qquad
%   \textsf{mov}~[r_m]~r_r \\
%   \hbox{(\TirNameStyle{WriteToRegFromMem})} \qquad 
%   \textsf{mov}~r_r~[r_m]
%   \end{array}
%   \]
Operationally, each first translates the virtual address stored in the register $r_m$ to a physical address \add{(a page table walk)},
then either updates the memory at that physical location with the contents of $r_r$ (for store)
or updates $r_r$ with the contents of that memory (for load).
% The semantics of each is realized by first translating the virtual-memory address of a memory location stored in $r_m$ to a physical memory location
% per the description in Section \ref{sec:backgroundonmachinemodel},
% then either updating those memory contents with the contents of register $r_r$ (\TirNameStyle{WriteToMemFromReg}) 
% or loading the value at that physical memory location into the register $r_r$
% (\TirNameStyle{WriteToRegFromMem}). 
Our formalization includes
additional \lstinline|mov| variants (e.g., accessing memory at constant offsets from the base register, or moves between registers),
basic integer and bitwise operations (\lstinline|add|,
\lstinline|and|, bit shifts, etc.) with their effects on \lstinline|rflags|, jumps and some (not all) conditional jumps, \lstinline|call|, \lstinline|ret|, \lstinline|push|,
and \lstinline|pop|.
The semantics model updates to the instruction pointer, though we elide that register (\lstinline|rip|)
from specifications in the paper for clarity.
\looseness=-1
